The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-08-24 adaptation/article/article.tex
80% of first draft done. Abstract missing, intro incomplete.
Changed figure 1 to now include the results from kim 2015.

\abstract{Abstract}
% \tableofcontents

\section{Introduction}
yeah

@@

\subsection{The model}
The model we present has three main components: (1) a context-inference component; (2) a motor adaptation component; (3) an action selection component. Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.

Based on the Bayesian brain hypothesis, we chose to cast both context inference and motor adaptation as Bayesian inference, in effect using a unified mechanism for the preparation of motor commands in new environments.

We now describe the three components of the model separately. We begin by describing the motor adaptation part for clarity, as the definition of ``context'' (used in context inference) is better described in terms of motor adaptation.

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to learn from observed errors. For example, we can learn to perform a reaching movement in front of us --a movement that any person has been practicing all their lives-- underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals. While a number of models for motor adaptation exist, we chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also does away with the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, let us return to reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \beta)
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the reach movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements.

As a day progresses, and especially if the day included a lot of physical activity of the arm, muscle exhaustion leads to a different response of the muscles in the arm to motor commands. However, we are still able to produce accurate reach movements. This happens through adaptation, where the observed error signals of a moment are used to temporarily update the parameters $\beta$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\beta | \vartheta, s, c) \propto p(\vartheta | s, c, \beta)p(\beta)
\end{equation}
where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$. $p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta$ is the previous estimate of $\beta$.

In this account of motor adaptation, the prediction error signal takes the form of an observed outcome $\vartheta$ that differs from that of the most likely outcome predicted by the forward model. The amount of adaptation that happens after observing an error will depend not only on the size of the error (how far away from the most likely outcome) it is, but also of how precise the existing estimate of $\beta$ is. We will discuss this further below.

Not only is the internal model for everyday movements not fixed (i.e. adaptation is possible), it is also not unique. Multiple internal models are used and updated, depending on the task at hand. For example, a person might have an internal model for baseline reaching movements (e.g. with empty hands) $f_B$, and a different one for reaching movements while holding a heavy object, $f_H$. A key aspect of motor adaptation is to select the relevant model to update, should a prediction error arise.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of hte environment (e.g. high winds when walking), as well as the task at hand. Previous models have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we focus on the idea that the context can be inferred by integrating several sources of information, an idea with great intuitive appeal. Through this idea, we explain many previously-unexplained phenomena from different experiments with humans.

Identifying any possible context with a categorical variable $\zeta$, the identification of a context can be done through Bayesian inference:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from instant to instant, and (2) an overall estimation of observing any one context, if the context did change. Below, we discuss how these two components effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions, but also any other sensory information that might be indirectly related to actions, or even completely intependent from them. For this section, the most important component of $\vartheta$ is contextual information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the contextual information can take the form of a visual cue [EGS], the place where the task must be carried out [EGS] or even the start of a new block of trials [EGS]. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1
}, c_{t-1})$ and are integrated together to identify a context.

@@

\section{Experimental results}
In this section, we go through a number of experimentally-observed phenomena and show how imperfect context inference, as done in our model, provides a parsimonious explanation for all of them.

In what follows, we show numerical results obtained the model outlined above. The full mathematical implementation of the model can be found in \sref{subsection:model-description}.

Before discussing these phenomena, let us introduce terminology. As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force. Depending on the trial (and experiment), the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point (approximately) in opposite directions. We represent contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g. clockwise), but the second one has half the magnitude. Finally, we represent error-clamp trials with the letter E and washout trials with W.

With this terminology, a typical experiment [EGs] would have a block structure of O-A-B-E (or O-A-(-A)-E), which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).

\subsection{Savings and quick de-adaptation}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants [A THOUSAND CITATIONS, c.f. papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}. In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of an adaptation rate apparently higher than that observed during the first A block.

@@

In this section, we will show that both phenomena are the two sides of a coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.

To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that is available. In some experiments [EGssss], the context is clearly \todo{Maybe not use the word 'cue', as many experiments don't have explicit sensory cues}cued to the participant, e.g. in the form of visual cues [EGs] or in changes to the location of the task [EGs]. We call these cued-context experiments. In other experiments, partial information is available to participants [EGsss]; we refer to these as partially-cued experiments. While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context. Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments. In this section, we focus on the two former categories, while the latter is left for the next sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands]. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance.

In FIGURE XXX, we show data from representative experiments from the cued-context \citep{Kim_Neural_2015} and partially-cued \citep[][group 1c]{Oh_Minimizing_2019} categories are shown. \cite{Kim_Neural_2015} used a colored light as a visual cue at the beginning of each experiment, whereas \cite{Oh_Minimizing_2019} included a condition in which the necessary adaptation was in the order of the motor error, making feedback-based context inference slow. It can be seen that in cued-context experiments \cite[data from][]{Oh_Minimizing_2019}, savings take the form of an immediate switch from the adaptation at trial $t$ to that in trial $t-1$, to the level that was learned before. In contrast, in partially-cued experiments, the switch is fast (compared to the initial adaptation), but not immediate. Our model explains this difference via context inference: in the cued-context experiments, context inference relies on the very reliable information of visual cues and the change in context is identified immediately; because of this, the model can re-activate the forward model adequate for the context from the first trial after the change occured. In contrast, in partially-cued experiments, the model needs to accumulate evidence in favor of the correct context throughout a number of trials, and the actions selected by the model are combinations of the actions that correspond to all the possible contexts (we discuss this in the following sections).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{./figures/figure_1.png}
\caption{Savings and de-adaptation. Data from our simulations (A and C) compared to data from \cite{Oh_Minimizing_2019} (B and D). In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number. The black lines represent the optimal adaptation, i.e. the size of the visuomotor rotation during the task. (A) A simulated experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements. It can be seen that after context switches (e.g. at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay. (B) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a. The same immediate switch can be observed as that in A. (C) Simulated experiment with a 10 degree adaptation. In this case, savings are much slower. (D) Data adapted from \cite{Oh_Minimizing_2019}, figure 4a.} \label{fig:oh-2019}
\end{figure}

Note that in the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model. \todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.

\section{The effects on the rate of adaptation}
In our model, adaptation is gated by the uncertainty on the current context. More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation, given a fixed size of prediction error.

@@

\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between two possibilities: no force (baseline O) a clockwise force (context A) and a counter-clockwise force (context -A). They divided participants into three groups for which the transition from one trial to the next were very uncommon, somewhat common and almost every trial. Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.

\cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning. The more volatile the environment, the less participants adapt their motor responses after observing an error. In \fref{fig:herzfeld-2014}, we show results obtained with our model in a paradigm matching that of \cite{Herzfeld_memory_2014}. We simulated the conditions of low volatility and maximum uncertainty ($z = 0.9$ and $z = 0.5$ in \cite{Herzfeld_memory_2014}) for thirty trials, repeating the experiment 300 times with random sequences of A and -A each time, following the transition probabilities from the experiment. In \fref{fig:herzfeld-2014}A we show the adaptation as a function of trial number, averaged across all 300 runs. It can be seen that the agent learns more slowly on the more volatile environment ($z = 0.5$) than on the less volatile one, matching the results from \cite{Herzfeld_memory_2014}. Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment.

@@

In our model, learning is gated by context inference: the higher the uncertainty on the context in which an error occured, the lower the effect of this error on future trials (i.e. adaptation).

To account for volatility, the model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods). Effectively, this transition term gives the model uncertainty on the current context, regardless of observations. The higher the volatility, the higher the uncertainty.

\section{Action selection}
As with learning, our model selects actions (motor commands) based on context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if some uncertainty exists, the selected action is influenced by all the possible current contexts, with a weight directly related to how likely each one of those contexts is.

Experimental evidence supporting this view can be found in experiments with context switching. For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group (where -A refers to an adaptation in the opposite direction as A, but with the same magnitude), with a block sequence A-(-A)-A-(-A), keeping the size of the necessary adaptation equal across both groups. After A and 3A in one group, A and -A in the other, had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A. The authors interpreted this as evidence that a new adaptation happens more quickly if it is in the same direction as the current adaptation (e.g. both counter-clockwise), and more slowly if they are in the opposite direction (e.g. clockwise to counter-clockwise).

Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context. As discussed above, action selection in our model is done via a weighted avarage:
\begin{equation}
p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
\end{equation}
In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out. Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A, as $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure by \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{./figures/figure_3.png}
\caption{Motor error when switching back to a previously learned adaptation. Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from +8 (yellow) or -8 (blue) to 4 is shown. (A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}. On the y axis, the error is shown in arbitrary units, related to centimiters through a monotonically increasing transformation with the same origin. (B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
\label{fig:davidson-2004}
\end{figure}

The model additionally predicts that the difference in observed switching speed between 3A-A and (-A)-A would disappear in an experiment in which the adaptations do not cross the baseline, i.e. all are positive or all are negative, or in an experiments in which the contexts are clearly cued. \todo{Maybe add simulations that show this, as well as simulations where the baseline context is not present.}

@@

\subsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials. However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment.

@@

\section{Discussion}
\begin{enumerate}
\item The effect of pauses?
\end{enumerate}

>>>>

\abstract{Abstract}
Very interesting abstract. Much wow.

\section{Introduction}
It has been long known that humans are capable of learning multiple novel motor tasks at the same time, even when these tasks oppose each other \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Functional_1997,}, although interference can occur between tasks \cite[e.g.][]{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}.

In order to study motor adaptation, motor error is defined as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \cite[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient is defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \cite{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \cite{Kim_Neural_2015} and saccade adaptation \cite{Catz_Cerebellardependent_2008}.

By introducing blocks of trials in which a manipulation is present (e.g. the mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error. A number of phenomena have been observed that persist across participants and across many different motor adaptation experiments: savings, quick de-adaptation, anterograde interference and spontaneous recovery.

In order to explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners [CITATION NEEDED], in which a single one-dimensional variable is used to describe the adaptation of motor commands given the observed motor errors. By having multiple time scales, each with a learning rate and a forgetting rate, these models have been fitted to experimental data, explaining these distinct experimental phenomena.

Bayesian accounts of motor adaptation have also been previously presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via prediction error generated by these internal models. These models replace the idea of explicit multiple time scales with an estimation of their own uncertainty: as the model practices the same task, the uncertainty on its estimate of the underlying modification is reduced, which lowers variability in responses, as well as increases the stability of the estimate.

While these general models of adaptation explain the most common phenomena observed in experiments, a number of known phenomena remain outside of the scope of these models. (1)

[unexplained phenomena. The truth is out there. I want to believe]

@@

\subsection{The model}
The model we present has three main components: (1) context-inference, (2) motor adaptation and (3) action selection. Each of these components informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For a complete mathematical description of the model, see the Methods section.

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement in front of us underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \cite{Wolpert_Multiple_1998a} \todo{Is this the original one? Include more}. While a number of models for motor adaptation exist [EGs], we chose to make use of an exact Bayesian learner. As we discuss below, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminates the need for explicit multiple time scales of learning. We show that this adaptation mechanism can deal with adaptation after any number of trials, ranging from a handful of trials in an experiment to the level of expertise a professional dancer has accrued through years of practicing the same movements.

To describe this component of the model, let us return to reaching movements. Throughout our lives, we have learned the equivalence between a motor command and its outcomes in the form of a forward model. We can write this forward model as follows:
\begin{equation}
p(\vartheta | s, c) = f(s, c; \beta)
\end{equation}
where $\vartheta$ are the outcomes of an action (in the example, direction and velocity of the reach movement), $s$ is the current state (e.g. the current position of the hand), $c$ is a motor command and $\beta$ are the parameters of the forward model $f(\cdot)$. Throughout our lives, the parameters $\beta$ have been fine tuned to produce an accurate forward model $f$ to guide our movements and it is these parameters which can be updated during motor adaptation.

When an observed outcome does not match the predictions produced by $f$, the brain uses the prediction error to update the parameters $\beta$ of the forward model. To describe this adaptation process, our model makes uses of Bayesian inference, such that:
\begin{equation}
q(\beta | \vartheta, s, c) \propto p(\vartheta | s, c, \beta)p(\beta)
\end{equation}
where $q(\beta | ...)$ represents the new estimate of $\beta$ after having observed the outcome $\vartheta$ of the previous motor command $c$. $p(\vartheta | s, c, \beta)$ is the \textit{a priori} probability of observing the outcome $\vartheta$ as predicted by the forward model $f$, and $p(\beta)$ is the previous estimate of $\beta$.

\subsection{Context inference}
In our model, the idea of context is quite general. The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking) and of the task at hand (immediate and long-term goals, reward structure, etc). Previous models \cite[e.g.][]{Wolpert_Multiple_1998a,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. $f_B$ and $f_H$ from the previous section) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context.

In this work, we expand on this idea of switching internal models by creating context inference that integrates several sources of information, including sensory cues, outcome prediction errors and reward prediction errors.  We show how the incorporation of all available contextual information explains many previously-unexplained phenomena from experiments with human subjects, without invoking \textit{ad hoc} model components.

Identifying any possible context with a categorical variable $\zeta$, the identification of a context is done by our model through Bayesian inference:
\begin{equation}
q(\zeta | \vartheta_t, \vartheta_{t-1}, c_{t-1}) \propto p(\vartheta_t | \vartheta_{t-1}, c_{t-1})p(\zeta)
\end{equation}
where the prior distribution over contexts $p(\zeta)$ includes information from two sources: (1) an expectation of continuity, encoding the expectation that the context does not change from trial to trial, and (2) an overall estimation of observing any one context, if the context did change. Below, we discuss how these two components effect different experimentally observed phenomena.

In the previous section, we used the variable $\vartheta$ to refer to the outcomes of actions. In this section, we expand the definition of an observation $\vartheta$ to include any information given to the decision-making agent by the environment. This includes, as before, the outcomes of actions (e.g. the new position of the hand), but also any other sensory information that might be indirectly related to actions, or even completely intependent from them. For this section, the most important component of $\vartheta$ is sensory contextual information, i.e. information that might directly help infer the identity of the current context.

In experimental settings, the sensory contextual information can take the form of a visual cue \cite[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \cite[e.g.][]{Forano_Timescales_2020} or even the start of a new block of trials \cite{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the priors $p(\vartheta$ form part of $p(\vartheta_t | \vartheta_{t-1}, c_{t-1})$ and are integrated together to identify a context.

By making use of NormalGamma priors on context, the model performs exact Bayesian inference at every trial. The exact update equations can be found in the Methods section.

@@

\section{Experimental results}
In this section, we go through a number of experimentally-observed phenomena and show that context inference, as done in our model, provides a parsimonious explanation for all of them.

For clarity, we introduce some terminology before discussing these phenomena. As an example, we will use a typical motor adaptation task in which participants have to make forward-backward reaching movements while holding the handle of a mechanical arm that exerts a force. Depending on the trial, the arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, it can be said that $B = (-A)$, as the forces point in opposite directions. We can also talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g. clockwise), but the second one has half the magnitude. Finally, error-clamp trials, in which the mechanical arm forces the participant to make straight-line movements, are represented with the letter E.

With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).

\subsection{Savings and quick de-adaptation}
Savings refers to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants [A THOUSAND CITATIONS, c.f. papers with volatility or ramping up/down], but animal experiments have shown mixed results \citep[e.g.][]{Kojima_Memory_2004}. In an O-A-O-A experiment, for example, savings would be seen in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

@@

In this section, we will show that both phenomena are the two sides of the same coin, and the speed at which this quick adaptation/de-adaptation happens, which is known not to be immediate [Citations] is related to the availability and reliability of contextual information, as well as to the quality of the prediction errors.

To show this, we categorized experiments performed with human participants by looking at the amount of contextual information that made available to participants. In some experiments [EGssss], the context is clearly revealed to the participant. We call these cued-context experiments. In other experiments, partial information is available to participants [EGsss] in the form of prediction errors, partial contextual information or reward prediction errors; we refer to these as partially-cued experiments. While there are no experiments in which no contextual information is available, there are some experiments which contain blocks of trials in which no contextual information is available and participants have no way of inferring the current context. Abusing our notation, we call these experiments (or rather, these sections of the experiments) uncued experiments. In this section, we focus on the two former categories, while the latter is left for the next sections.

Note that, because context inference integrates information from many sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information [EGs of experiments with hidden hands]. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than normal trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite[e.g.][]{Ethier_Spontaneous_2008}.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results is quite evident, both in the immediacy of the switches as well as in the fact that learning continues from the state it was when the context was last observed.

\begin{figure}
\centering
\includegraphics{./figures/figure_1.png}
\caption{Savings and de-adaptation. Data from our simulations (left column) compared to data from \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} (right column). In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number. The black lines represent the optimal adaptation, i.e. the size of the visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015} with two rotations (-40 and 40 degrees), in addition to baseline. The switches between contexts were visually cued and therefore switches are immediate, as can be seen in both simulations (left) and the experimental data from \cite{Kim_Neural_2015} (right). (B) An experiment in which the agent must adapt to a 20 degree visuomotor rotation during shooting movements. It can be seen in both the simulations (left) and experimental data from \cite{Oh_Minimizing_2019} (right) that after context switches (e.g. at trial 120), the agent returns to the adaptation of the previous block, with one or two trials of delay. (C) Same experiment as (B) but  with a 10 degree adaptation. In this case, savings are much slower.} \label{fig:oh-2019}
\end{figure}

\cite{Oh_Minimizing_2019} performed a similar experiment in which contexts were not visually cued. In addition, the authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model. \cite{Oh_Minimizing_2019} immediate switching in the first experiment, while switching was much slower in the second experiment. As can be seen on the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large, and slow, noisy switches when it is low. This difference is explained by our model in terms of the size of the adaptation: the smaller the adaptation is, the more difficult it is for the agent to notice when the context has changed. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

Note that the data from \cite{Oh_Minimizing_2019} and \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model. \todo{Need to be careful here, as de-adaptation is sometimes discussed only in terms of experiments with no contextual cues or where the meaning of the cues is inverted}.

\section{The effects on the rate of adaptation}
In our model, adaptation is gated by the uncertainty on the current context. More specifically, a prediction error observed at trial $t$ will effect motor adaptation with a magnitude proportional to the size of the error, where the proportionality constant is related to the uncertainty over the context in which the error was observed: the higher the uncertainty, the lower the size of the adaptation.

@@

\cite{Herzfeld_memory_2014} used an experimental paradigm in which the force that a mechanical arm exerts on the participant's hand can change from trial to trial between O, A and -A. They divided participants into three groups for which the context transition from one trial to the next was very uncommon, somewhat common and almost every trial. Importantly, these transitions were uncued, and the participant could only infer that a transition had occurred after a motor error was observed.

\cite{Herzfeld_memory_2014} found that the volatility of the environment affects learning. The more volatile the environment, the less participants adapt their motor responses after observing an error. In \fref{fig:herzfeld-2014}, we show results obtained with our model in a paradigm matching that of \cite{Herzfeld_memory_2014}. We simulated the conditions of low volatility and maximum uncertainty ($z = 0.9$ and $z = 0.5$ in \cite{Herzfeld_memory_2014}) for thirty trials, repeating the experiment 300 times with random sequences of A and -A each time, following the transition probabilities from the experiment. In \fref{fig:herzfeld-2014}A we show the adaptation as a function of trial number, averaged across all 300 runs. It can be seen that the agent learns more slowly on the more volatile environment ($z = 0.5$) than on the less volatile one, matching the results from \cite{Herzfeld_memory_2014}. Additionally, \fref{fig:herzfeld-2014}B shows the sensitivity to errors, here defined as $\text{adaptation} / \text{error}$; the more volatile environment creates a lower sensitivity to errors than the predictable environment, a difference that is consistent across all trials.

@@

To account for volatility, our model includes a term in context inference which assumes a transition from the previous context to all other possible contexts (see Methods). Effectively, this transition term gives the model uncertainty on the current context, regardless of observations. The higher the volatility, the higher the uncertainty. This higher uncertainty is what lowers the observed learning rate across all trials.

\section{Action selection}
As with learning, our model selects actions (motor commands) based on context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is.

Experimental evidence supporting this view can be found in experiments with context switching. For example, \cite{Davidson_Scaling_2004} reported an experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A). After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A. The authors interpreted this as evidence that switching between adaptations happens more quickly if it is in the same direction as the current adaptation (e.g. both counter-clockwise), and more slowly if they are in the opposite direction (e.g. clockwise to counter-clockwise).

Our model offers a different explanation for the observed results: the assymetry is due to the existance of the baseline context. As discussed above, action selection in our model is done via a weighted avarage:
\begin{equation}
p(a_t | s_t) \propto \displaystyle \sum_{m \in M} p(m | s_t, s_{t-1}, a_{t-1}) p(a_t | m)
\end{equation}
In experiments without cued contexts, the baseline model $m_O$ has a non-zero probability $p(m | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out. Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, essentially ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but on the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -8 context shifts back to 4 more slowly than the one exposed to +8, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B.

\begin{figure}
\centering
\includegraphics{./figures/figure_3.png}
\caption{Motor error when switching back to a previously learned adaptation. Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from +8 (yellow) or -8 (blue) to 4 is shown. (A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}. On the y axis, the error is shown in arbitrary units, related to centimiters through a monotonically increasing transformation with the same origin. (B) Data adapted from \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
\label{fig:davidson-2004}
\end{figure}

@@

\subsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be described as the display of spontaneous recovery (i.e. displaying adaptation consistent with a previously-encountered context) during the early trials of the E block, followed by a slow return to baseline across as many as hundreds of trials. However, the direction of the spontaneous recovery, its length, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment [EGs].

@@
